Geometric models reveal the hidden structure of conceptual knowledge

### 2023-01-22 paper/main.tex
added knowledge map figure, finished a complete draft of the methods section (excluding the "estimating held-out knowledge" section, which is waiting on our remaining analyses-- see slack discussion for more info)

% supplemental figures

@@

@@

\begin{equation}
    \hat{k}(x) = \frac{\sum_{i \in \mathrm{correct}} \mathrm{RBF}(x, q_i, \lambda)}{\sum_{j = 1}^N \mathrm{RBF}(x, q_j, \lambda)},
\end{equation}
where
\begin{equation}
    \mathrm{RBF}(x, \mu, \lambda) = \exp\left\{-\frac{||x - \mu||^2}{\lambda}\right\}.
\end{equation}

>>>>

% supplemental figures
\newcommand{\individualKnowledgeMapsA}{S1}
\newcommand{\individualKnowledgeMapsB}{S2}
\newcommand{\individualKnowledgeMapsC}{S3}
\newcommand{\individualLearningMapsA}{S4}
\newcommand{\individualLearningMapsB}{S5}

@@

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{figs/knowledge_and_learning_maps}

    \caption{\textbf{Mapping out the geometry of knowledge and learning.}
    \textbf{A. Average ``knowledge maps'' estimated using each quiz.} Each map
    displays a 2D projection of the estimated knowledge about the content
    reflected by \textit{all} regions of topic space (see \textit{Creating
    knowledge and learning map visualizations}). The topic trajectories of each
    lecture and the coordinates of each question are indicated by dotted lines
    and dots. Each map reflects an average across all participants. For
    individual participants' maps, see
    Figures~\individualKnowledgeMapsA,~\individualKnowledgeMapsB,
    and~\individualKnowledgeMapsC. \textbf{B. Average ``learning maps''
    estiamted between each successive pair of quizzes.} The learning maps are
    in the same general format as the knowledge maps in Panel A, but each
    coordinate in the learning maps indicates the \textit{difference} between
    the corresponding coordinates in the indicated \textit{pair} of knowledge
    maps-- i.e., how much the estimated knowledge ``changed'' across the two
    quizzes. Each map reflects an average across all participants. For
    individual participants' maps, see
    Figures~\individualLearningMapsA~and~\individualLearningMapsB. \textbf{C.
    Word clouds for sampled points in topic space.} Each word cloud displays
    the relative weights of each word reflected by the blend of topics
    represented at the locations of the stars in the maps. The words' colors
    indicate how much each word is weighted on average across all timepoints'
    topic vectors in the \textit{Four Fundamental Forces} (blue) and
    \textit{Birth of Stars} (green) videos, respectively.}

    \label{fig:knowledge-maps}
    \end{figure}

@@

An important feature of our approach is that, given a trained text embedding
model and participants' quiz performance on each question, we can estimate
their knowledge about \textit{any} content expressable by the embedding model--
not solely the content explicitly probed by the quiz questions. To visualize
these estimates
(Figs.~\ref{fig:knowledge-maps},~\individualKnowledgeMapsA,~\individualKnowledgeMapsB,~\individualKnowledgeMapsC,~\individualLearningMapsA,
and~\individualLearningMapsB), we used UMAP~\citep{McInEtal18a} to define a 2D
projection of the text embedding space. Sampling the original 100-dimensional
space at high resolution to obtain an adequate set of topic vectors spanning
the embedding space would be computationally intractable. However, sampling a
2D grid is much more feasible. We defined a rectangle enclosing the 2D
projections of the lectures' and quizzes' embeddings, and we sampled points
from a regular $100 \times 100$ grid of coordinates that evenly tiled the
enclosing rectangle. We sought to estimate participants' knoweldge (and
learning-- i.e., changes in knowledge) at each of the resulting 10000
coordinates.

To generate our estimates, we placed a set of 39 radial basis functions (RBFs)
throughout the embedding space, centered on the 2D projections for each
question (i.e., we included one RBF for each question). At coordinate $x$, the
value of an RBF centered on a question's coordinate $\mu$, is given by:
\begin{equation}
    \mathrm{RBF}(x, \mu, \lambda) = \exp\left\{-\frac{||x - \mu||^2}{\lambda}\right\}.
    \label{eqn:rbf}
\end{equation}
The $\lambda$ term in the RBF equation controls the ``smoothness'' of the
function, where larger values of $\lambda$ result in smoother maps. In our
implementation we used $\lambda = 50$.  Next, we estimated the ``knowledge''
at each coordinate, $x$, using:
\begin{equation}
    \hat{k}(x) = \frac{\sum_{i \in \mathrm{correct}} \mathrm{RBF}(x, q_i, \lambda)}{\sum_{j = 1}^N \mathrm{RBF}(x, q_j, \lambda)}.
    \label{eqn:rbf-knowledge}
\end{equation}
Intuitively, Equation~\ref{eqn:rbf-knowledge} computes the weighted proportion of
correctly answered questions, where the weights are given by how nearby (in the 2D space)
each question is to the $x$.  We also defined \textit{learning maps} as the coordinate-by-coordinate
differences between any pair of knowledge maps.  Intuitively, learning maps reflect the \textit{change}
in knowledge across two maps.
